---
title: "Effectivness of skin cancer drugs by analyzing meta data"
subtitle: "STA130 - Final Project"
author: "Umi Yamaguchi, Ralisha Woodhouse, Yifei Hong"
date: December 8, 2022
output: 
  beamer_presentation:
    theme: "Szeged"
    colortheme: "beaver"
    fonttheme: "professionalfonts"
    slide_level: 2
classoption: "aspectratio=169"
fontsize: 8pt
urlcolor: cyan
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Introduction

\Large 

### Goals And Motivation 

\vspace{.5cm}
1.

\vspace{.5cm}
2.

\vspace{.5cm}
3.

### Population and Samples

```{r, size= 0.5} 
"Hello, World."
```
* Samples


## Data Wrangling 

\large

### Setup
* Population
* Samples
* Population
* Samples

## Checking Outliers and Data Sets

\large

*post the heat maps/box plots etc...

## Two Hypothesis Teting 

\large

### Questions: 

\vspace{.5cm}
* 

### Methods: 

\vspace{.5cm}
1.


## Results and Prediction (Two Hypothesis Testing)

\large

### Results
*Talk about the results
*


## Linear Regressions 

\large

### Questions: 

\vspace{.5cm}
* 

### Methods: 

\vspace{.5cm}
1.


## Results and Prediction (Linear Regressions)

\large

### Results
*Talk about the results
*

## Classification Trees 

\large

### Questions: 

\vspace{.5cm}
* 

### Methods: 

\vspace{.5cm}
1.


## Results and Prediction (Classification Trees)

\large

### Results
*Talk about the results
*

## Overall Prediction and Results

### Prediction
1.
2.
3.

### Results
*
*
*

## Limitation
*
*
*
*
*

## Conclusion with Future Perspective

### Conclusions
*
*
*

### How this results will help?
1.
2.
3.











